---
title: 'Reproducible Research: Peer Assessment 1'
output:
  pdf_document: default
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data
1. Load the data (i.e. `read.csv()`)
```{r}
unzip(zipfile="activity.zip")
data <- read.csv("activity.csv", sep = ",")
steps <- as.numeric(data$steps, na.rm = TRUE)
```
## What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day
```{r}
plotHist <- hist(steps)
plotHist
```
2. Calculate and report the **mean** and **median** total number of steps taken per day
```{r}
print(summary(steps))
```
## What is the average daily activity pattern?
1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
library(ggplot2) #load the ggplot2 library
avg <- aggregate(list(data$steps), list(data$interval), FUN = mean, na.rm = TRUE) #set the mean for each interval in the dataset and assign it to "avg"
colnames(avg) <- c("interval", "steps") # reset the column names
plot <- ggplot(avg, aes(x = interval, y = steps)) # set the data for the plot using ggplot to the variable "plot"
plot + geom_step(direction = "hv") + xlab("Interval") + ylab("Average Steps") + ggtitle("Average Steps at each interval")
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
avg[which.max(avg$steps),]
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
# ----- Imputting missing values -----
stepsNa <- mean(is.na(data$steps)) # (1) percentage of missing(NA) values in the dataset
print(stepsNa * 100)
```
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

There are only NA values in the steps observation. 
The plan is to create a new data frame and replace the NA's values with the mean of steps.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
# ----- Create dataset with mean steps filled in for NA values -----
data2 <- data # (3)
meanSteps <- as.integer(mean(data$steps, na.rm = TRUE))
data2$steps[which(is.na(data2$steps))] <- meanSteps
```
4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
plot <- ggplot(data2, aes(steps))
plot + geom_histogram() + ggtitle("Histogram of steps - replacing NA's with mean")
print(summary(data2$steps))
```
## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
data2$date <- as.Date(data2$date, "%Y-%m-%d")
data2$daysWk <- weekdays(data2$date)
data2$daysWk <- ifelse(data2$daysWk == "Saturday" | data2$daysWk == "Sunday", "Weekend", "Weekday")
```
2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
ggplot(data2, aes(interval, steps)) + facet_grid(daysWk ~ .) + geom_line()
```